---
title: "Polling Simulation: Sampling and the Central Limit Theorem"
---

In this activity, you'll use data from the 2024 Cooperative Election Study (CES) — a large-scale national survey conducted during the 2024 election — to simulate what happens when a news organization tries to estimate public opinion from a poll sample. The full dataset has nearly 55,000 respondents, which we'll treat as our population.

Your goal is to see firsthand how sample size affects the accuracy of estimates, and what the distribution of sample estimates actually looks like — the core idea behind the Central Limit Theorem.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```

### Task 2: Load the CES data

Load the CES 2024 religion dataset from this URL and save it to a dataframe called `ces`:

`https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/ces_2024_religion.csv`

```{r}
ces <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/ces_2024_religion.csv")
```

### Task 3: Calculate the population proportion

We'll focus on one question: what share of American adults identify as religiously unaffiliated (atheist, agnostic, or nothing in particular)? This has been a major story in American religion journalism for years.

First, we'll create a new column called `unaffiliated` that is `TRUE` if a respondent's `trad` value is `"Atheist"`, `"Agnostic"`, or `"Nothing in Particular"`, and `FALSE` otherwise. Then calculate the proportion of the full dataset that is unaffiliated. Save this value as `population_proportion`.

```{r}
ces <- ces |>
  mutate(unaffiliated = trad %in% c("Atheist", "Agnostic", "Nothing in Particular"))

proportion <- mean(ces$unaffiliated)
```

That population proportion is the "true" population value that any poll is trying to estimate. Write it down — you'll be comparing everything else to it.

### Task 4: Take a single small sample

Use `set.seed(42)` and then sample 50 rows from `ces`. Save it as `sample_50`. Calculate the proportion of unaffiliated respondents in your sample.

```{r}
set.seed(42)
sample_50 <- ces |> sample_n(50)
mean(sample_50$unaffiliated)
  

```

How close is your sample estimate to the original proportion? Is it surprising?

ANSWER HERE

### Task 5: Take a single larger sample

Now sample 500 rows. Set a different seed value, grab the sample and save it as `sample_500`. Calculate the proportion of unaffiliated respondents in this sample.

```{r}
set.seed(405)
sample_500 <- ces |> sample_n(500)
mean(sample_500$unaffiliated)

```

How did the estimate change compared to the sample of 50? What does this suggest about sample size?

ANSWER HERE

### Task 6: Simulate many polls

A single sample can get lucky or unlucky. To understand how polling actually works, we need to see the *distribution* of estimates across many samples. The code below draws 1,000 samples of size 50, calculates the proportion of unaffiliated respondents in each, and stores the results.

```{r}
set.seed(3)

simulate_polls <- function(data, sample_size, n_simulations = 1000) {
  proportions <- replicate(n_simulations, {
    s <- data |> sample_n(sample_size)
    mean(s$unaffiliated)
  })
  tibble(proportion = proportions, sample_size = sample_size)
}

simulations_50 <- simulate_polls(ces, 50)
simulations_500 <- simulate_polls(ces, 500)
```

### Task 7: Visualize the simulation results

Create two histograms side by side — one for the 1,000 samples of size 50 and one for the 1,000 samples of size 500. Add a vertical line at `proportion` to each so you can see how well the samples are centering on the true value.

```{r}
simulations_all <- bind_rows(simulations_50, simulations_500) |> 
  mutate(sample_size = paste("n =", sample_size))

ggplot(simulations_all, aes(x = proportion)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  geom_vline(xintercept = proportion, color = "red", linetype = "dashed", linewidth = 1) +
  facet_wrap(~sample_size) +
  labs(
    title = "Distribution of Sample Estimates: Share Religiously Unaffiliated",
    subtitle = "Red line = true population proportion. Each histogram shows 1,000 simulated polls.",
    x = "Sample Proportion",
    y = "Count",
    caption = "Source: 2024 Cooperative Election Study"
  ) +
  theme_minimal()
```

### Task 8: Reflect on what you see

**What shape do both histograms have?** Describe them in plain language. Does it surprise you that this shape emerges from a variable that is simply TRUE or FALSE for each respondent?

ANSWER HERE
Both of them have a triangular shape, somewhat similar normal distribution with no skews. The shapes make sense, it seems that the n=500 just has more to go off so it looks more complete. Since there are only two values, there shouldn't be anything too crazy that sticks out as we can see.

**Compare the two histograms.** What is the most visible difference between the n=50 and n=500 distributions? What does that difference tell you about what happens when a poll uses a larger sample?

ANSWER HERE
The most visible difference is the space in between the bins. That tells us that we just need more volume, more values and people to test or data to input in order for us to really draw a complete picture. If you put both diagrams in a story, readers might look at the n=50 as less than or pretty pointless just based off the unused space and lack of volume when it comes to the data we took down. In a way when you use less you feel like you're leaving out part of the story. 

**Both distributions are centered near the red line.** What does it mean that 1,000 different random samples — each one drawn independently — tend to cluster around the same value? What would it mean for journalism if that *weren't* the case? It means that the data at least in this situation is straight forward. There can only be so much difference when the population's subjects stay the same you just change the amount. If this wasn't the case then surveys and certain data populations would be pretty much useless if you couldn't mirror certain stories or data points. If you did this poll in different states or different times in our country's history, the data would look different. 

ANSWER HERE

**A reporter tells you their story is based on a poll of 50 people.** Based on what you've seen in the simulation, what questions would you want to ask before trusting that number? You don't need to use any technical terms — just describe what you observed.
I would want to ask about the 50 people and how diverse they were. Ages, races occupations all matter. You can't get a complete picture of really any group with just 50 people unless it's a group of 50 or 100 people. It also depends on what we are trying to gather, if it is yes or no questions for college students, I might just say get to 75 or 100 since it's low lift but also it can help to further and paint/complete the picture. Of course 50 could work and if we're on a time crunch it could work, but we just want more variance and more data, it's not like it hurts. 

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
